A sample workflow

### 2025-04-29 analyses/african_mammals_workflow.qmd
Fixing a lot of typos and removing superlatives (and -- eek! -- this was in two paragraphs, so many more exist)

We expect scaling up to more countries, species and populations may cause us to think carefully about the data we have in hand versus the question we aimed to answer. While many recent papers treat the LPI data as effectively a well stratified and randomized sample or even effectively a good cencus, it is clearly not. Thus, models that try to deal with such uneven sampling and build in the more of the structure of data, would be far more appropriate. For example, most models for election forecasts now include regularization via partial pooling, post-stratificatiob and other methods designed to adjust for the fact that many polls are 'non-representative.' Such polls may be more similar to LPI data. Our model includes some regularization via partial pooling, but we may need far more of an observational model added to our population demography model to accommodate the uneven sampling. Estimates from such a model may be far more uncertain, but likely also more correct given our combination of question and data.

This may lead us to realize we really don't have great data for this question and we would then need to prioritize better data collection efforts if we truly want good estimates to answer this question.

@@

The small workflow we should here may seem like extra work to those who have used it, and it is compared to many ecological workflows in practice. We argue that the extra effort is worth it given the critical questions we are addressing, and also that the process becomes much quicker with practice. The first time may be slow, but the resulting model will be better for it (even if it is still arguably simple, as the one we show here is), but the following times will go faster with likely improved models.

>>>>

We expect scaling up to more countries, species and populations may cause us to think carefully about the data we have in hand versus the question we aimed to answer. While many recent papers treat the LPI data as (effectively) a well stratified and randomized sample or even a good census, the LPI data is clearly neither. Thus, models that try to deal with such uneven sampling and build in the more of the structure of data, would be more appropriate. For example, most models for election forecasts now include regularization via partial pooling, post-stratification and other methods designed to adjust for the fact that many polls are 'non-representative.' Such non-representative polls may be very similar to LPI data. Our model includes some regularization via partial pooling, but we may need much more of an observational model added to our population demography model to accommodate the uneven sampling. Estimates from such a model may be more uncertain, but likely also more correct given our combination of question and data.

This may lead us to realize we really don't have great data for this question and we would then need to prioritize better data collection efforts if we want good estimates to answer this question.

@@

The small workflow we show here may seem like extra work to those who have used it, and it is compared to many ecological workflows in practice. We argue that the extra effort is worth it given the critical questions we are addressing, and also that the process becomes much quicker with practice. The first time may be slow, but the resulting model will be better for it (even if it is still arguably simple, as the one we show here is), but the following uses of this iterative workflow will go faster with likely improved models.
